---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fastpipe

This package proposes an alternative to the pipe from the *magrittr* package. 
It's named the same and passes all the *magrittr* test so can easily be a drop-in
replacement, its main advantages is that it is faster and solves most of the issues
of *magrittr*.

Install with :

``` r
remotes::install_github("fastpipe")
```
## Families of pipes and performance


* A `%>%` family of pipes reproduce and extends the features offered by *magrittr*
* A `%>>%` family of pipes behaves the same, but doesn't support functional sequences (`. %>% foo()`) nor compound assignment (`bar %<>% baz()`)
* A `%>>>%` family of pipes behaves the same but doesn't support implicit dots

For the 2 latter, skipping these steps has a dramatic effect on performance. We provide
a benchmark below :

```{r}
library(fastpipe)
`%.%` <- fastpipe::`%>%` # (Note that a *fastpipe*, unlike a *magrittr* pipe, can be copied)
`%>%` <- magrittr::`%>%`
bench::mark(
  'magrittr::`%>%`' =
    1 %>% identity %>% identity() %>% (identity) %>% {identity(.)},
  'fastpipe::`%>%`' =
    1 %.% identity %.% identity() %.% (identity) %.% {identity(.)},
  'fastpipe::`%>>%`' =
    1 %>>% identity %>>% identity() %>>% (identity) %>>% {identity(.)},
  'fastpipe::`%>>>%`' =
    1 %>>>% identity(.) %>>>% identity(.) %>>>% identity(.) %>>>% identity(.),
  base = identity(identity(identity(identity(1))))
)
rm(`%>%`) # reseting `%>%` to fastpipe::`%>%`
```

We see that our standard `%>%` pipe is twice faster, and that our `%>>>%` pipe
is 10 times faster (on a properly formatted call).

## Issues solved and special features

*fastpipe* passes all *magrittr*'s tests, but it doesn't stop there and solves
most of its issues, at least most of the open issues on GitHub.

* Lazy evaluation

```{r}
# https://github.com/tidyverse/magrittr/issues/195
# Issue with lazy evaluation #195
gen <- function(x) {function() eval(quote(x))}
identical(
  {fn <- gen(1); fn()},
  {fn <- 1 %>% gen(); fn()})
```

``` {r, eval = FALSE}
# https://github.com/tidyverse/magrittr/issues/159
# Pipes of functions have broken environments
identical(
  {
    compose <- function(f, g) { function(x) g(f(x)) }
    plus1   <- function(x) x + 1
    compose(plus1, plus1)(5)},
  {
    plus2 <- plus1 %>% compose(plus1)
    plus2(5)
})
#> [1] TRUE
# Note : copy and pasted because didn't work with knitr
```

* It considers `!!!.` as `.` when deciding wether to insert a dot

This was requested by Lionel and seems fairly reasonable as its is unlikely that
a user will nedd to use both `.` and `!!!.` in a call.

```{r}
letters[1:3] %>% rlang::list2(!!!.)
```

* `:::`, `::` and `$` get a special treatment to solve a common issue

```{r}
iris %>% base::dim
iris %>% base:::dim
x <- list(y = dim)
iris %>% x$y
```

* It fails explicitly in some cases rather than allowing strange behavior silently

```{r, error = TRUE}
iris %>% head %<>% dim
. %<>% head
```


* The new pipe `%S>%` allows the use of *rlang*'s `!!!` operator to splice dots in
any function.

```{r}
c(a = 1, b = 2) %S>% data.frame(!!!.)
```

* The new pipe `%L>%` behaves like `%>%` except that it logs to the console the 
calls and the time they took to run.

```{r}
1000000 %L>% rnorm() %L>% sapply(cos) %>% max
```

## Implementation

The implementation is completely different from *magrittr*, we can sum it up as
follow : 

* `%>>>%` pipes : evaluate *rhs* in parent environment, overriding `.` with the value
  of the *lhs*. Its code is extremely straightforward

```{r}
`%>>>%`
```

* `%>>%` pipes : same as above but use heuristics to insert dots, so `x %>% head` becomes
   `x %>% head(.)` etc
   
```{r}
`%>>%`
```  
   
* `%>%` pipes : same as above but must support functionnal sequences and compound assignment,
  we do so by modifying global variables when we find a relevant start
  to the pipe chain and reseting them when going out.

*fastpipe* operators contain their own code while in *magrittr*'s current
implementation they all have the main code and are recognized in the function
by their name. Moreover the pipe names are not hardcoded in the functions, which
prevents confusion like the fast that `%$%` still sometimes work in *magrittr*
when only `%<%` is reexported. https://github.com/tidyverse/magrittr/issues/194

The pipes have a class, the functional sequences don't, they are regular functions
using the `%>>>%` family of *fastpipes*.

```{r}
`%T>>%`
. %>% sin %>% cos() %T>% tan(.)
```

## Benchmarks for functional sequences

An interesting and little known fact is that using functional sequences in functionals
is much more efficient than using lambda functions calling pipes, for instance :
`purrr::map(foo, ~ .x %>% bar %>% baz)` will often be much slower than
`purrr::map(foo, . %>% bar %>% baz)`. I tried to keep this nice feature but
didn't succeed to make it as efficient as in *magrittr*. The difference show only
for large number of iteration and will probably be negligible in any realistic loop.

```{r}
`%>%` <- magrittr::`%>%`
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# functional sequence with 5 iterations
walk <- purrr::walk
bench::mark(
  purrr_lambda =
  walk(1:5, ~.x %>% identity %>% identity() %>% (identity) %>% {identity(.)}),
  magrittr =
  walk(1:5, . %>% identity %>% identity() %>% (identity) %>% {identity(.)}),
  fastpipe =
  walk(1:5, . %.% identity %.% identity() %.% (identity) %.% {identity(.)})
)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# functional sequence with 1000 iterations
bench::mark(
  purrr_lambda =
    walk(1:1000, ~.x %>% identity %>% identity() %>% (identity) %>% {identity(.)}),
  magrittr =
    walk(1:1000, . %>% identity %>% identity() %>% (identity) %>% {identity(.)}),
  fastpipe =
    walk(1:1000, . %.% identity %.% identity() %.% (identity) %.% {identity(.)})
)
```


## A few notes 

This is an experiment, and might still change change.

The current package might be well suited for programming because it's robust,
fast, and very light.

